---
title: "CS100 Studio 5"
author: "Mariel Pacada"
date: "10/14/2019"
output: html_document
---

### Swine Flu Data Cleaning

```{r swine_setup, include = FALSE}
swineflu <- read.csv("https://cs.brown.edu/courses/cs100/studios/data/5/swineflu_mod.csv")
colnames(swineflu)[colnames(swineflu) == "Vaccination.rates.per.1000"] <- "Vaccine.pmp"
library(dplyr)
library(ggplot2)
```

```{r clean_cases}
levels(swineflu$Cases)
levels(swineflu$Cases)[levels(swineflu$Cases) == "four"] <- "4"
levels(swineflu$Cases)

newCases <- as.numeric(as.character(swineflu$Cases))
class(newCases)

swineflu$Cases <- newCases

range(swineflu$Cases)
```

```{r hist}
hist(swineflu$Cases, xlab = "Confirmed Cases", main = "Distribution of Cases")
```

```{r fix_range}
filter(swineflu, swineflu$Cases < 0)
swineflu$Cases[which(swineflu$Cases < 0)] <- 20
```

```{r new_hist}
hist(swineflu$Cases, xlab = "Confirmed Cases", main = "Distribution of Cases")
```

```{r fix_country}
swineflu$Country <- as.character(swineflu$Country)
```

```{r fix_na}
swineflu <- na.omit(swineflu)
```

```{r save_file}
filename_csv <- paste("cleanSwineflu", Sys.Date(), ".csv", sep = "")
write.csv(swineflu, file = filename_csv)

filename_rds <- paste("cleanSwineflu", Sys.Date(), ".rds", sep = "")
saveRDS(swineflu, file = filename_rds)
```

&nbsp;

### Preliminary Analysis

```{r cases}
top_ten_cases <- swineflu %>%
                   select(Country, Cases) %>%
                   arrange(desc(Cases))

head(top_ten_cases, 10)
```
Above are the first ten countries that had the most cases of the swine flu. The US was below only to China.

```{r infect}
top_ten_infect <- swineflu %>%
                    select(Country, Infection.rate.pmp) %>%
                    arrange(desc(Infection.rate.pmp))

head(top_ten_infect, 10)
```

```{r death}
top_ten_death <- swineflu %>%
                   select(Country, Deaths.pmp) %>%
                   arrange(desc(Deaths.pmp))

head(top_ten_death, 10)
```

```{r vaccine}
top_ten_vaccine <- swineflu %>%
                     select(Country, Vaccine.pmp) %>%
                     arrange(desc(Vaccine.pmp))

head(top_ten_vaccine, 10)
```
The United States was the most vaccinated country, and, in turn, it was not in the top ten for either highest death rates or highest infection rates. 

```{r cor_case_death}
ggplot(swineflu, aes(x = Cases, y = Deaths)) + geom_point()
cor(swineflu$Cases, swineflu$Deaths)
```

```{r cor_infect_death}
ggplot(swineflu, aes(x = Infection.rate.pmp, y = Deaths.pmp)) + geom_point()
cor(swineflu$Infection.rate.pmp, swineflu$Deaths.pmp)
```
Intuitively, number of deaths and number of cases are highly correlated -- the higher the recorded number of cases, the higher such of deaths. However, we find that the infection rate and death rate are not at all strongly correlated. 

```{r cor_vac_infect}
qplot(swineflu$Vaccine.pmp, swineflu$Infection.rate.pmp)
cor(swineflu$Vaccine.pmp, swineflu$Infection.rate.pmp)
```
One would expect that the more people get vaccinated in a country, the less people get infected. However, this plot and the correlation coefficient show that these two variables are not strongly correlated. This means that a country's vaccination rate cannot predict its infection rate. 

```{r infect_outliers}
swineflu$Country[which(swineflu$Vaccine.pmp > 250)]
swineflu[swineflu$Country == "Australia",]

swineflu <- subset(swineflu, swineflu$Country != "Australia")
qplot(swineflu$Vaccine.pmp, swineflu$Infection.rate.pmp)
cor(swineflu$Vaccine.pmp, swineflu$Infection.rate.pmp)
```
The correlation is even closer to 0 than with the outlier. 

```{r death_outliers}
qplot(swineflu$Vaccine.pmp, swineflu$Deaths.pmp)
cor(swineflu$Vaccine.pmp, swineflu$Deaths.pmp)

swineflu$Country[which(swineflu$Vaccine.pmp > 400)]
swineflu <- subset(swineflu, swineflu$Country != "Korea (Republic of)")
swineflu <- subset(swineflu, swineflu$Country != "United States of America")

cor(swineflu$Vaccine.pmp, swineflu$Deaths.pmp)
```
The correlation is closer to what one would expect. 

### HDI Data Cleaning

```{r merge_setup, include = FALSE}
swineflu <- read.csv("https://cs.brown.edu/courses/cs100/studios/data/5/swineflu_mod.csv",
                     stringsAsFactors = FALSE)
hdi <- read.csv("https://cs.brown.edu/courses/cs100/lectures/scripts/section1/2014hdi.csv",
                stringsAsFactors = FALSE)
continents <- read.csv("http://cs.brown.edu/courses/cs100/lectures/scripts/section1/continents.csv", stringsAsFactors = FALSE)
```

```{r clean_hdg}
hdg <- as.factor(hdi$Human.Development.Group)

table(hdi$Human.Development.Group)
table(hdg)

hdi$Human.Development.Group <- hdg
```

```{r combine}
combined <- merge(swineflu, hdi, by = "Country")
swineflu[which(!(swineflu$Country %in% hdi$Country)), ]$Country
```

```{r fix_country}
swineflu$Country[which(swineflu$Country == "United States of America")] <- "United States"
swineflu$Country[which(swineflu$Country == "Bosnia and Herzegovinia")] <- "Bosnia and Herzegovina"
swineflu$Country[which(swineflu$Country == "Canada ")] <- "Canada"
swineflu$Country[which(swineflu$Country == "Uruguay ")] <- "Uruguay"
swineflu$Country[which(swineflu$Country == "Bolivia")] <- "Bolivia (Plurinational State of)"
swineflu$Country[which(swineflu$Country == "Iran")] <- "Iran (Islamic Republic of)"
swineflu$Country[which(swineflu$Country == "Laos")] <- "Lao People's Democratic Republic"
swineflu$Country[which(swineflu$Country == "Venezuela")] <- "Venezuela (Bolivarian Republic of)"
swineflu$Country[which(swineflu$Country == "Vietnam")] <- "Viet Nam"

swineflu[which(!(swineflu$Country %in% hdi$Country)), ]$Country
```

```{r clean_combine}
clean_combined <- merge(swineflu, hdi, by = "Country")
```

```{r combine_continents}
colnames(continents)[colnames(continents) == "name"] <- "Country"

continents$Country[which(continents$Country == "United States of America")] <- "United States"
continents$Country[which(continents$Country == "United Kingdom of Great Britain and Northern Ireland")] <- "United Kingdom"

continents_combined <- merge(continents, clean_combined, by = "Country")
```

### Further Analysis

```{r}



```