---
title: "CS100 Studio 8"
author: "Mariel Pacada"
date: "11/5/2019"
output: html_document
---

### Part 1A
```{r plot}
x <- runif(50, 0, 100)
y <- runif(50, 0, 100)

plot(x, y)
cor(x, y)
```
There seems to be no correlation between x and y. The correlation coefficient confirms this.

```{r regression}
my_model <- lm(y ~ x)
plot(x, y) + abline(my_model)
summary(my_model)
plot(my_model, which = 1)
```
The slope of the linear model is -0.013, which is insignificant because its p-value is almost 1. The R-squared value is 0.0002, which means that the model does not explain the data very well.

(Note: The slope, p-e, and R-squared values may be different at the time of knit because x and y are randomly generated. Nonetheless, the numbers should still be similar.)

### Part 1B
```{r plot2}
x <- runif(50, 0, 100)
e <- runif(50, -10, 100)
y <- 2 + 4 * x + e

my_model <- lm(y ~ x)

cor(y, x)
plot(x, y)
lm(y ~ x)
```
The scatterplot for x and y seems to indicate that the correlation between the two are close to positive 1. The slope and intercept predicted by the linear model are similar to the true values. 

```{r regression2}
plot(x, y) + abline(my_model, col = "red") + abline(a = 2, b = 4, col = "blue") + legend("topleft", legend = c("Best Fit", "True Model"), col = c("red", "blue"), lty = 1)

plot(lm(y ~ x), which = 1)
````


